---
title: MDS setup check, R
author: UBC Master of Data Science
# knitr is Quarto's R engine. There are no Python chunks in this document, so no
# reticulate is involved and nothing has to locate this project's Python from
# inside R. Python is checked by check-quarto.qmd instead.
engine: knitr
format:
  pdf:
    output-file: check-quarto-r-latex.pdf
  typst:
    output-file: check-quarto-r-typst.pdf
  html:
    output-file: check-quarto-r.html
---

## What this file checks

Rendering this document exercises the R half of the toolchain through Quarto:
Quarto has to find R, run the code below, and hand the result to a typesetter.
That is a different path from `check-rmarkdown.Rmd`, which R Markdown renders
itself.

```{r}
R.version$version.string
```

```{r}
data.frame(
  course   = c("DSCI 511", "DSCI 521", "DSCI 523"),
  block    = c(1, 1, 2),
  language = c("Python", "Python", "R")
)
```

## Characters that are not plain English

Montréal · naïve · Öl · 5 °C · α β γ · 10 – 20 · "curly quotes" · ✅ ❌ 📊 ⚠️

Mathematics is typeset separately from ordinary text, so it works everywhere:
$\alpha = 0.05$ and $\hat{\beta} = (X^{\top}X)^{-1}X^{\top}y$.

**What you should see.** In the HTML and in the Typst PDF, every character above
appears. In the LaTeX PDF the accents, the dashes and the mathematics appear, but
the Greek letters and the emoji do not — LaTeX has no glyph for them and drops
them silently. If an assignment needs emoji in a PDF, render it with Typst.
