---
title: "DSCI 521 Lab 0b"
subtitle: Submitting an RMarkdown Assignment
output:
  pdf_document:
    toc: yes
    toc_depth: '4'
---

This assignment is similar to the `lab0a` assignment
that used Jupyter notebooks.
In this lab assignment,
we will practice working with RMarkdown assignments.

## RMarkdown Documents

This is an RMarkdown document.
A literate programming document where you can interweave
prose text, code, and code output.

### Prose Text

The main difference from an RMarkdown document and a
a Jupyter notebook is that an RMarkdown document is easier to read
in any plain text editor and does not need any special tools
to render the document in a human readable way.

The other difference is that you can write regular prose text
in the document directly,
you do not need separate markdown cells.
You will only need specially formatted code chunk to run code.

In MDS you will primarily work with Jupyter notebooks for your Python assignments,
and you will use RMarkdown documents for your R assignments.

### Code Chunks

Below is a code cell that uses the `R` engine to run R code.

The code cell will have a play button for you to run the code.

```{r}
print("Hell, World!")
```

Your IDE will most likely highlight the code chunk in a different color,
so you can visually see where the code chunk area is.
If you accidently change or delete a character that specifies the beginning
and end of the code chunk,
the background coloring of your document may change,
and you may not see the code runing controls that you'd expect.

You can run the code by selecting the code and and pressing CTRL + Return (Windows/Linux),
or command/apple + Return (MacOS).

```{r}
print("Hell, Again!")
```

Optional: There's a package called `emo` with the function `ji()`
that you can install and use text to print emoji's.

The fundamental building block of R code is an expression.
Code chunks can contain multiple lines with multiple expressions.
When you run a chunk,
the lines of code are executed in the order in which they appear.
Every `print()` expression prints a line.
Run the next cell and notice the order of the output.

```{r}
print("First this line is printed,")
print("then this one,")
```

### Comments

When a line of **code** begins with the `#` character,
it is called a comment,
all the code and text to the right of a leading `#` character will
be ignored by the programming language.



Change the cell below by replacing the comment such that the output
will print the following results:

```
First this line is printed,
then this one,
and then the whole world.

```

*Note:* The setup for RMarkdown for PDF exporting requires a bit more
setup than we want to deal with right now.

```{r}
# your code here
```

## Writing RMarkdown files

RMarkdown is a a special file based on the "Markdown" syntax.
You can think of it as giving Markdown documents the ability
to specify a programming language and run code.

You can even run Python code in an RMarkdown document,
but you will need and set up an R libray called `reticulate` to do that.

Markdown documents are ubiquitous in data science and computing in general.

To add text, you type directly in the document where you want the text to be.
To add a new code chunk you can manually type in the markdown verbatim
fences with the `{r}` that specifies the engine.
In Rstudio you can also look for the green `+C` button up top to have it
enter a code chunk for you.

Below: add a code cell from scratch that prints out:

```
A Whole new chunk!
```

## Errors

R is a language,
and like natural human languages,
it has rules.
It differs from natural language in two important ways:

1. The rules are *simple*.
   You can learn most of them in a few weeks and
   gain reasonable proficiency with the language in a semester.
2. The rules are *rigid*.
   If you're proficient in a natural language,
   you can understand a non-proficient speaker,
   glossing over small mistakes.
   A computer running Python code is not smart enough to do that.

Whenever you write code, you'll make mistakes.
When you run a code cell that has errors,
Python will sometimes produce error messages to tell you what you did wrong.

Errors are okay; even experienced programmers make **many** errors.
When you make an error,
you just have to find the source of the problem, fix it, and move on.

We have made an error in the next cell.  Run it and see what happens.

```{r error=TRUE}
# un-comment the code below and fix it
#print("This line is missing something."
```

The error message will show
`Error: Incomplete expression: ` or ` unexpected end of input`.
If we look at the code, we will notice that we are missing the closing `)`

The `error=TRUE` in the code chunk is an option that allows the document to
render the error (more on that at the very end of the lab).

Fix the code above so it no longer errors.

## Stop, Restart, and Re-Run Chunks

You can stop and restart the R session by going to the `Session` drop down
in the RStudio IDE.

As you work,
you will typically want to run `Session` > `Restart R and Run All Chunks`
on a regular basis, so you can check that you did not introduce any
code errors as you work and edit.

> Everything in this file is editable, it's really easy to change a character
> somewhere that will break your code.
> Really, use the Restart R and Run All Chunks action often.
> You don't want to finish an assignment, and realize it won't render anymore.

## Autograded Questions

Many assignments in MDS will be distributed as files like this one.

You may need the following libraries.
Run the following code that checks if you have the `testthat` and `digest` packages,
and it will install it for you if it's missing.

```{r}
if (!require(testthat)) {install.packages("testthat")}
if (!require(digest)) {install.packages("digest")}
```

An autograded question will look like the one below:


Create object named `y` with a value of `10`.

```{r tags=c("otter_assign_solution_cell")}
# BEGIN SOLUTION
y <- 10
# END SOLUTION

print(y)
```

```{r}
. = ottr::check("tests/Q2-1-autograder.R")
```

Now, at the top of the RStudio IDE,
click the blue `Knit` button, and it should create a PDF file
of your assignment.
You will use this `.pdf` file as one of your homework submissions.
