---
title: "DSCI 521 Lab 0b: your first MDS R assignment"
author: "Your name here"
format:
  typst:
    output-file: lab0b.pdf
  html:
    output-file: lab0b.html
---

This is `lab0a` again, in R. Everything it taught applies here -- the same autograder,
the same export routes, the same rule about a document that has to run top to bottom.
What changes is the file: an **R Markdown** document rather than a notebook.

The difference worth knowing on day one is that this file is **plain text**. Prose is
written directly, with no cell around it, and code goes in a *chunk* that starts and
ends with a code fence:

```{r}
#| echo: fenced

# your code here
```

Open it in a text editor and it still
reads as a document, which is not true of an `.ipynb`.

Work top to bottom, running each chunk with its ▶ button or `Ctrl+Shift+Enter`.

```{r setup}
# ggplot2 is part of the tidyverse, which your R courses are built on and which the MDS
# setup check installs for you. ottr is the autograder.
library(ggplot2)
library(ottr)
```

## 1. Running code, and reading what comes back

A chunk either **prints** something or **returns** something. `print()` writes a line
as the code runs. The value of the last expression is shown underneath, with no
`print()` involved.

```{r}
print("Hello, MDS!")
print("This line is printed second.")
```

```{r}
7 * 6
```

Only the **last** expression is shown. This chunk computes two things and you see one
of them.

```{r}
3 + 3
7 * 6
```

## 2. A result with structure

Most results in this program are tables. R calls one a `data.frame`.

```{r}
courses <- data.frame(
  course = c("DSCI 511", "DSCI 521", "DSCI 523", "DSCI 531", "DSCI 552"),
  block = c(1, 1, 2, 2, 3),
  language = c("Python", "", "R", "R, Python", "R")
)
courses
```

## 3. A figure

Assignments ask for plots, and a plot has to survive being rendered into whatever you
hand in. This one is drawn with **ggplot2**, which is what your R courses use: you name
the data, say which column goes on which axis inside `aes()`, then add the layers you
want with `+`.

```{r}
ggplot(courses, aes(x = factor(block))) +
  geom_bar() +
  labs(x = "block", y = "courses", title = "Courses per block")
```

## 4. Answering in your own words

Not every question is code. Some ask you to explain something, and in an R Markdown
document you answer by typing straight into the file -- no cell to create, no mode to
switch.

### Question 4.1

Look at the figure above. **In two or three sentences, say which block has the most
courses and what that means for how busy you will be.** Replace the line below.

_Your answer here._

## 5. Mathematics -- read the source of this file

Assignments in the statistics and machine learning courses are full of equations. You
will mostly *read* them rather than write them, but you should recognise the two forms.

Inline, inside a sentence: the mean of $n$ observations is $\bar{x}$, reported with a
standard deviation $s$.

On its own line, centred:

$$
\bar{x} = \frac{1}{n} \sum_{i=1}^{n} x_i
$$

**Open this file in a text editor.** One dollar sign either side puts maths in a
sentence; two put it on its own line. It is the same syntax as `lab0a` -- nothing about
maths changes between Python and R, because neither language is involved.

Greek letters are written as commands: `$\alpha$` gives $\alpha$, `$\beta$` gives
$\beta$, `$\sigma^2$` gives $\sigma^2$. Write them that way rather than pasting the
character and they survive every export format.

## 6. Tables -- read the source of this too

| block | courses | mostly |
| --- | --- | --- |
| 1 | 2 | Python |
| 2 | 2 | R |
| 3 | 1 | R |

The columns need not line up in the source; the dashes on the second line are what
make it a table.

## 7. Emoji, and why your PDF may lose them

You can put emoji in text 🎉 and print them from code. Run the chunk below.

```{r}
print("Kernel busy: \U0001F535")
print("All tests passed: \U00002705")
print("Look at the plot \U0001F4CA")
```

Now the part worth knowing. **Whether those emoji appear depends on how you export this
document**, and when they vanish nothing warns you -- the sentence around them stays
perfectly readable:

| export route | emoji | Greek as `$\alpha$` | Greek pasted as a character |
| --- | --- | --- | --- |
| HTML | yes | yes | yes |
| PDF via Typst | yes | yes | yes |
| PDF via LaTeX | **no** | yes | **no** |

`make pdf` in this folder uses **Typst**, which is why that is the command you are
given. RStudio's `Knit` button takes the LaTeX route instead, which is worth knowing
before you press it and wonder where the emoji went.

## 8. Questions the autograder marks

You write code, run `ottr::check(...)`, and it tells you straight away whether you were
right. Run it as often as you like -- it is feedback, not a submission.


### Question 8.1

Count how many of the five courses use R at all -- note that one of them uses both
languages, and one lists none. Assign that number to `r_courses`. **You have not
learned any R yet, so the answer is written in the chunk below** -- this question is
about the autograder, not about the code.

**Do the wrong thing first, on purpose.** Type `r_courses <- 2`, run the chunk, then
run the check underneath it. Read what a failure looks like. *Then* put in the real
answer and run both again. The autograder disagreeing with you is the normal state of
affairs, not a problem.

```{r}
# The `language` column from the table above, as a plain vector.
languages <- c("Python", "", "R", "R, Python", "R")

# You have not written any R yet, so here is the answer. Replace the NULL below with:
#
#     sum(grepl("R", languages))

r_courses <- sum(grepl("R", languages))  # SOLUTION

r_courses
```

```{r}
. = ottr::check("tests/q1.R")
```

### Question 8.2

Most of what you write in this program will be **functions**. Write one called
`block_size` that takes a vector of block numbers and one block number, and returns how
many times that block appears. The answer is again given in the chunk.

> **This question has hidden tests.** The check runs some of its cases in front of you
> and keeps the rest back until your work is marked. Look at how many cases it reports,
> and remember that the number you see is not the number that count. Most MDS
> assignments work this way: passing the checks you can see is necessary, not
> sufficient.

```{r}
# Again, the answer. Replace the NULL with the line below it:
#
#     sum(blocks == block)

block_size <- function(blocks, block) {
  # BEGIN SOLUTION
  sum(blocks == block)
  # END SOLUTION
}

# Works on a plain vector and on a column of the table alike.
block_size(courses$block, 1)
```

```{r}
. = ottr::check("tests/q2.R")
```

## 9. Break something on purpose

Everyone writes broken code, constantly, and the only skill that matters is reading
what R says about it. So make an error deliberately, look at it, then put it back.

### Question 9.1

The chunk below works. **Delete the closing bracket `)` from the end of the line and
run it.** Read the message.

```{r}
print("Nothing is wrong with this line yet.")
```

You should see something ending in:

```
Error in parse(...) : unexpected end of input
```

R stops at the first thing it cannot read and tells you where it gave up. That is not
always where the mistake is -- a bracket left open earlier makes R complain about the
*next* line -- so read upwards from the error as well as at it.

**Now put the bracket back and run the chunk again.** Leave it working: the next
section renders the whole document from a clean start, and a broken chunk stops it.

A document you hand in has to run top to bottom without errors. A chunk you broke and
forgot about is the most common reason an assignment that "worked" fails when a TA
opens it.

## 10. Handing it in

```bash
make all
```

That renders `lab0b.pdf` and `lab0b.html`, re-running every chunk from a clean start.
Submit the PDF.

Run it before you finish. If it fails, your document only worked because of something
you ran earlier and then changed -- which is exactly what section 9 was about.
